Interpretable Machine Learning

### 2019-01-14 chunks/story-wines.Rmd
continues talk

##  {.emphasizedabit}
You want to predict wine quality from its physicochemical properties.

@@

## Step 1: Find data
Free dataset of red and white "Vinho Verde" wine.

```{r}
knitr::include_graphics("../images/minho.png")
```

Three sensory assessors (using blind tastes), which graded the wine in a scale that ranges from
0 (very bad) to 10 (excellent)

@@

<!-- Draw: -->
<!-- Volatile acidity as polish remover, pH as pH paper, alchohol as liqour bottle, chlorides as salt prinkler, citric acid as lemon,  -->
<!-- sulfur as mushroom?,  -->

## Step 1: Find data

@@

# Step 2: Throw ML on your data

##

Compare different models with 10x CV
- Linear regression model
- Decision tree
- Random forest

@@

## Step 2: Throw ML on your data

@@

bmr = benchmark(lrns, tsk, rdesc, measures = list(mae))
bmr_tab = getBMRAggrPerformances(bmr, as.df = TRUE)
knitr::kable(bmr_tab[-1])
```

=> The random forest (ranger) is the best model.

@@

## Step 2: Throw ML on your data

@@

## {data-background=../images/delivery.gif data-background-size=contain}

@@

##  {.emphasizedabit}

Client: "We would love to learn some insights."

@@

TODO: Slide explaining permutation feature importance

@@

TODO: Slide explaining ALE plots

## How do features affect predictions?
Method: Accumulated Local Effects
```{r}
eff = FeatureEffect$new(pred, "alcohol", method = 'ale')
plot(eff)
```

## How do features affect predictions?
Method: Accumulated Local Effects
```{r}
eff = FeatureEffect$new(pred, "volatile.acidity", method = 'ale')
plot(eff)
```

@@

## Interactions between alcohol and volatile acidity?

@@

## Rule of thumb for wine quality?

```{r surrogate}
library(partykit)
tree = TreeSurrogate$new(pred, maxdepth  = 2)
node_inner2 = node_inner(tree$tree, pval = FALSE, id=FALSE)
node_terminal2 = node_boxplot(tree$tree, id = FALSE)
plot(tree$tree, inner_panel = node_inner2, terminal_panel = node_terminal2)

@@

The tree explains `r sprintf("%.2f%s", 100 *  rsquared, "%")` of the black box prediction variance.

## Exceptionally bad wine

@@

TODO: Image of really bad wine

Predicted quality: `r predictions[which(predictions == min_pred),]`

@@

pred$predict(worst2())[[1]]

worst3 = function(){
  w = worst
  w$alcohol = 13
  w$volatile.acidity = 1
  w
}

pred$predict(worst3())[[1]]
```

>>>>

##  {.emphasizedabit}
Predict wine quality from its physicochemical properties.

@@

# Step 1: Find data

## {.center data-background=../images/minho.png data-background-size=contain}

## Wine Dataset

- ~6500 red and white Portuguese "Vinho Verde" wines
- Features: physiochemical properties
- Quality assessed by blind tasting, from 0 (very bad) to 10 (excellent)

@@

## Quality Distribution

@@

# Step 2: Apply Machine Learning

## Compare Models with 10x CV {.emphasizedabit}

- Linear regression
- Decision tree
- Random forest

@@

@@

bmr = benchmark(lrns, tsk, rdesc, measures = list(mae))
bmr_tab = getBMRAggrPerformances(bmr, as.df = TRUE)
#knitr::kable(bmr_tab[-1])
```

@@

## Compare Models with 10x CV {.emphasizedabit}

- Linear regression (`r sprintf('%.2f', bmr_tab$mae.test.mean[bmr_tab$learner.id == "regr.lm"])`)
- Decision tree (`r sprintf('%.2f', bmr_tab$mae.test.mean[bmr_tab$learner.id == "regr.rpart"])`)
- Random forest (`r sprintf('%.2f', bmr_tab$mae.test.mean[bmr_tab$learner.id == "regr.ranger"])`)

=> The random forest is the best model.

## Prediction vs. Actual Quality

@@

@@

##  {.emphasizedabit .center data-background=../images/no-idea.gif data-background-size=cover}

@@

```{r}
knitr::include_graphics("../images/permimp-shuffle.png")
```

@@

```{r}
knitr::include_graphics("../images/pdp-rent.png")
```s

## How do features affect predictions?
```{r}
eff = FeatureEffect$new(pred, "alcohol", method = 'ale')
plot(eff)
```

## How do features affect predictions?
```{r}
eff = FeatureEffect$new(pred, "volatile.acidity", method = 'ale')
plot(eff)
```

@@

## Interaction: Alcohol + volatile acidity

@@

## Rule of thumb for wine quality? {.center data-background=../images/global-surrogate.png data-background-size=contain}

## Rule of thumb for wine quality? 
```{r surrogate}
library(partykit)
tree = TreeSurrogate$new(pred, maxdepth  = 2)
node_inner2 = node_inner(tree$tree, pval = FALSE, id=FALSE)
node_terminal2 = node_boxplot(tree$tree, id = FALSE)
plot(tree$tree, inner_panel = node_inner2, terminal_panel = node_terminal2)

@@

Tree explains `r sprintf("%.2f%s", 100 *  rsquared, "%")` of black box prediction variance.

## Explain individual prediction (e.g. worst wine)

@@

##  {.center data-background=../images/bad-wine.gif data-background-size=contain}

@@

worst3 = function(){
  w = worst
  w$alcohol = 13
  w$volatile.acidity = 1
  w
}
```

### 2019-01-14 chunks/tools.Rmd
continues talk

## Model-specific methods

@@

## Model-agnostic methods

@@

<div class="notes">
- Partial dependence plots
- LIME
- ...
</div>

@@

## Model-agnostic methods: Global Surrogate {.center data-background=../images/global-surrogate.png data-background-size=contain}

## Model-agnostic methods: Local Surrogate {.center data-background=../images/local-surrogate.png data-background-size=contain}

>>>>

## Model-specific Methods

@@

@@

@@

## Model-agnostic Methods: Global Surrogate {.center data-background=../images/global-surrogate.png data-background-size=contain}

## Model-agnostic Methods: Local Surrogate {.center data-background=../images/local-surrogate.png data-background-size=contain}
